---
title: "Chicago crime reports"
author: "Rafal Nowicki"
date: "15 04 2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(bigrquery)
library(rgdal)
library(sp)
```
Setting up the project:
```{r}
project <- "rafal-mannheim"
```
Setting up the connection:
```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project)

con
```
Available tables:
```{r}
dbListTables(con)
```
Data description [here](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2)
How the data looks like:
```{sql connection= con}
SELECT *
FROM crime
WHERE year = 2015
LIMIT 50
```
How many reports are there?
```{sql connection= con}
SELECT COUNT(*) num_cases
FROM crime
WHERE year = 2015
```
Select the example data:
```{r}
crime <- tbl(con, "crime")

crime_data <- crime %>%
              select(year, community_area, arrest) %>%
              filter(year == 2015 & arrest == TRUE) %>%
              group_by(community_area) %>%
              summarise(arrest_num = n()) %>%
              collect()

crime_data_stand <- crime_data %>%
                    mutate_at(2, scale)
```
Shape files for plotting maps of Chicago are available on the following website:

https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6

As we downloaded required files let's unpack it and make use of them.
Now we can see how community areas around Chicago look like.
```{r message=FALSE"}
map <- readOGR(".", "chicago_shape")
map <- spTransform(map, "+proj=longlat")
```
Let's merge the data with map:
```{r}
spatial_data <- merge(map, crime_data_stand, by.y = "community_area", by.x = "area_num_1")
```
Plot it:
```{r}
pal <- colorRampPalette(c("steelblue","white" ,"darkred"), bias = 2.5)

arrest_plot <- spplot(spatial_data, zcol = "arrest_num", colorkey = TRUE, col.regions = pal(100), cuts = 99,
                   par.settings = list(axis.line = list(col =  'transparent')),
                   main = "Number of arrest in Chicago, 2015")

arrest_plot
```
Now lower level data - chicago beats
[here](https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Beats-current-/aerh-rz74)

Map:
```{r message=FALSE}
map_beat <- readOGR(".", "chicago_beats")
map_beat <- spTransform(map_beat, "+proj=longlat")
```
Data:
```{r}
crime <- tbl(con, "crime")

crime_data_beat <- crime %>%
              select(year, beat, arrest) %>%
              filter(year == 2015 & arrest == TRUE) %>%
              group_by(beat) %>%
              summarise(arrest_num = n()) %>%
              collect()

crime_data_beat_stand <- crime_data_beat %>%
                    mutate_at(2, scale)
```
Let's merge the data with map:
```{r}
spatial_data_beat <- merge(map_beat, crime_data_beat_stand, by.y = "beat", by.x = "beat_num")
```
Plot it:
```{r}
arrest_plot_b <- spplot(spatial_data_beat, zcol = "arrest_num", colorkey = TRUE, col.regions = pal(100), cuts = 99,
                   par.settings = list(axis.line = list(col =  'transparent')),
                   main = "Number of arrest in Chicago, 2015")

arrest_plot_b
```
